DRAFT \\ From physical principles to classical and quantum \\ Hamiltonian and Lagrangian particle mechanics

### 2016-09-26 2015CompleteDerivation/completeDerivation.tex
Adding quantum inner product and projections

To be on par with classical mechanics we need at least address three problems. We need to be able to compare two quantum states and ask what fraction of them share the same configuration: this will give us an inner product. We need to be able to express state variable distributions and expectations: this will give us linear operators. We need to identify the conjugate quantities $k_i$ and $p_i$. And we need to be mindful that the observations made after \ref{prop:differentiable_manifold} for the  classical case are still valid: $\Psi$ cannot be a complete metric space. Let's start with the first.

@@

From the norm and the projection we can define a product $< \cdot |\cdot> : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\mathsf{P}_{\psi_2} (\psi_1) \equiv \frac{<\psi_2|\psi_1>}{|\psi_2|^2} \psi_2$. If the arguments are the same, the projection returns the same state, therefore $<\psi_1|\psi_1> = |\psi_1|^2$: the product is positive definite. Since the projection is linear, the product is linear as well.

The coefficient $\alpha$ is a function of both $\psi_1$ and $\psi_2$, so we write it as $\alpha = <\psi_2|\psi_1>$. It is linear in $\psi_1$ as $\mathsf{P}_{\psi_2}$ is linear. The square module $|<\psi_2|\psi_1>|^2$ gives us the amount of material in $\psi_1$ that has the same distribution of $\psi_2$. $|<\psi_1|\psi_1>|^2$ As all states

Suppose we are .

@@

@@

\begin{justification}
	We claim the vector field $S \in T\mathcal{Q}$ admits a potential $H$ such that $S^{a} \omega_{ab} = \partial_{b}H$. Let $v, w \in T_m M$ be two vectors defined at a point $m \in M$. Let $v^a, w^b$ their components. Let $v'\equiv f v, w' \in T_{f(m)}M$ be the pushforward of $v, w$ by $f$. Since $f$ is a symplectomorphism we have $v^{a} \omega_{ab} w^{b} = v'^{a} \omega_{ab} w'^{b}$. The vector components change according to $v'^a = \partial_b \xi(t+dt)^a v^b = (\delta^a_b + \partial_b S^a dt) v^b$. We have:
	\begin{align*}
	v^{a} \omega_{ab} w^{b} &= v'^{a} \omega_{ab} w'^{b}  \\
	&= (v^{a} + \partial_{c} S^{a} v^{c} dt) \omega_{ab} ( w^{b} + \partial_{d} S^{b} w^{d} dt) \\
	&= v^{a} \omega_{ab} w^{b} + (\partial_{c} S^{a} v^{c} \omega_{ab} w^{b} \\
	&+ v^{a} \omega_{ab} \partial_{d} S^{b} w^{d}) dt + O(dt^2)
	\end{align*}
	$v^{c} w^{b} \partial_{c} S_{b} - v^{a} w^{d} \partial_{d} S_{a} = 0$ where $S_{b} \equiv S^{a} \omega_{ab}$. The relationship must true for and pair of vector, therefore $\partial_{a} S_{b} - \partial_{b} S_{a} = curl(S_{a}) = 0$. $S$ is a curl free vector field and admits a potential $H$ such that $S_{a} = \partial_{a}H$.
\end{justification}

>>>>

To be on par with classical mechanics we need at least address three problems. We need to be able to compare two quantum states and ask what fraction of them share the same configuration: this will give us an inner product. We need to be able to express state variable distributions and expectations: this will give us linear operators. We need to identify the conjugate quantities $k_i$ and $p_i$. Let's start with the first.

@@

This projection has other useful properties. If $\psi_1$ is already prepared like $\psi_2$, then its projection on $\psi_2$ will be equal to $\psi_1$. Conversely, if the projection of $\psi_1$ is equal to itself, then it means that the state is already prepared like $\psi_2$. That is: $\mathsf{P}_{\psi_2} (\psi_1) = \psi_1 \Leftrightarrow \psi_1 = \alpha \psi_2$. Moreover, because the material is irreducible, if $23\%$ of $\psi_1$ is prepared like $\psi_2$, then $23\%$ of $\psi_2$ is prepared like $\psi_1$. If not, we could learn more (or less) about $\psi_1$ by studying the part of $\psi_2$ that is prepared like $\psi_1$. Therefore we have $\frac{|\mathsf{P}_{\psi_2} (\psi_1)|^2}{|\psi_1|^2}=\frac{|\mathsf{P}_{\psi_1} (\psi_2)|^2}{|\psi_2|^2}$. Lastly, if we take the projection of $\psi_1$ along $\psi_2$ and we project it again on $\psi_1$, we end up with a part of $\psi_1$. As this is a proper part of $\psi_1$, its state is just $\psi_1$ rescaled, no change in correlation. That is: $\mathsf{P}_{\psi_1} (\mathsf{P}_{\psi_2} (\psi_1)) = a \psi_1$ with $a \in \mathbbm{R}$.

From the norm and the projection we can define a complex product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\mathsf{P}_{\psi_2} (\psi_1) \equiv \frac{\langle \psi_2|\psi_1 \rangle}{|\psi_2|^2} \psi_2$. Because of the properties of the norm and the projections, the product is positive definite, linear and conjugate symmetric. This means the state space $\Psi$ for an irreducible material is an inner product space.

\begin{prop}\label{prop:inner_product}
	The state space $\Psi$ for an irreducible material is an inner product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$.
\end{prop}
\begin{justification}
	TODO
\end{justification}

Note that the inner product only has an indirect physical definition, based on the norm and the projections. We'll therefore want to base our physical intuition on the latter concepts.

Also note that the observations made after \ref{prop:differentiable_manifold} for the  classical case are still valid. $\Psi$ cannot be a complete metric space as it does not include discontinuous functions. Therefore it is only a pre-Hilbert space. For mathematical convenience, one can take its completion with respect to the norm induced by the inner product and obtain a Hilbert space. Yet, such object does no longer represent a physically meaningful entity.

As such, we need to make sure that the rest of our definitions and justifications do not require a complete metric space to be valid or they would not be physically meaningful. In particular, we cannot write the usual link $\psi(q) = \langle q | \psi \rangle$ between the wave function and the inner product as the material cannot be prepared at a specific value if $q$ is a continuous variable.

%If the arguments are the same, the projection returns the same state, therefore $<\psi_1|\psi_1> = |\psi_1|^2$: the product is positive definite. Since the projection is linear, the product is linear as well.

%The coefficient $\alpha$ is a function of both $\psi_1$ and $\psi_2$, so we write it as $\alpha = <\psi_2|\psi_1>$. It is linear in $\psi_1$ as $\mathsf{P}_{\psi_2}$ is linear. The square module $|<\psi_2|\psi_1>|^2$ gives us the amount of material in $\psi_1$ that has the same distribution of $\psi_2$. $|<\psi_1|\psi_1>|^2$ As all states

@@

\begin{thrm}\label{thrm:function_topology}
	Let $X$ and $Y$ be Hausdorff and second countable topological spaces. Let $C(X, Y) \equiv \{f: X \rightarrow Y \; | \; f$ is continuous$\}$ be the space of continuous functions endowed with the Compact-Open Topology $\mathsf{T}$. Then $C(X, Y)$ is a Hausdorff and second countable topological space.
\end{thrm}

\begin{proof}
	We claim $C(X, Y)$ is Haudorff. Let $f, g \in C(X, Y)$ such that $f \neq g$. $\exists x \in X$ such that $f(x)\neq g(x)$. Let $K = \{x\}$. $K$ is compact.
\end{proof}

@@

\begin{proof}
	We claim the vector field $S \in T\mathcal{Q}$ admits a potential $H$ such that $S^{a} \omega_{ab} = \partial_{b}H$. Let $v, w \in T_m M$ be two vectors defined at a point $m \in M$. Let $v^a, w^b$ their components. Let $v'\equiv f v, w' \in T_{f(m)}M$ be the pushforward of $v, w$ by $f$. Since $f$ is a symplectomorphism we have $v^{a} \omega_{ab} w^{b} = v'^{a} \omega_{ab} w'^{b}$. The vector components change according to $v'^a = \partial_b \xi(t+dt)^a v^b = (\delta^a_b + \partial_b S^a dt) v^b$. We have:
	\begin{align*}
	v^{a} \omega_{ab} w^{b} &= v'^{a} \omega_{ab} w'^{b}  \\
	&= (v^{a} + \partial_{c} S^{a} v^{c} dt) \omega_{ab} ( w^{b} + \partial_{d} S^{b} w^{d} dt) \\
	&= v^{a} \omega_{ab} w^{b} + (\partial_{c} S^{a} v^{c} \omega_{ab} w^{b} \\
	&+ v^{a} \omega_{ab} \partial_{d} S^{b} w^{d}) dt + O(dt^2)
	\end{align*}
	$v^{c} w^{b} \partial_{c} S_{b} - v^{a} w^{d} \partial_{d} S_{a} = 0$ where $S_{b} \equiv S^{a} \omega_{ab}$. The relationship must true for and pair of vector, therefore $\partial_{a} S_{b} - \partial_{b} S_{a} = curl(S_{a}) = 0$. $S$ is a curl free vector field and admits a potential $H$ such that $S_{a} = \partial_{a}H$.
\end{proof}

\begin{thrm}
	Scalar product
\end{thrm}

\begin{proof}
	We claim $\langle x , y \rangle = \langle x, \mathsf{P}_x(y)\rangle$. Let $x, y \in X$. $\langle x , y \rangle x = |x|^2 \mathsf{P}_x(y) = |x|^2 \mathsf{P}_x(\mathsf{P}_x(y)) = \langle x, \mathsf{P}_x(y)\rangle$.

	We claim $\mathsf{P}_x(y) = a x$ for some $a \in \mathbbm{C}$. Let $x, y \in X$. $\mathsf{P}_x(\mathsf{P}_x(y)) = \mathsf{P}_x(y)$. $\mathsf{P}_x(y) = a x$ for some $a \in \mathbbm{C}$

	We claim $\langle a x , y \rangle = a^\dagger \langle x , y \rangle$. Let $x, y \in X$ and $a \in \mathbbm{C}$. $\langle a x , y \rangle = \langle a x, \mathsf{P}_x(y)\rangle = \langle a x, b x\rangle$ for some $b \in \mathbbm{C}$. $\langle a x, b x\rangle = \frac{b}{a} \langle a x, a x\rangle = \frac{b}{a} |a x|^2 = \frac{b}{a} |a|^2 |x|^2 = \frac{a^\dagger ab}{a} \langle x, x\rangle = a^\dagger \langle x, b x\rangle = a^\dagger \langle x, \mathsf{P}_x(y)\rangle = a^\dagger \langle x , y \rangle$
\end{proof}
